---
title: "Laboratory 1 - Introduction to R"
author: Tamy Tsujimoto and Victor Ritter
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output: 
  html_document:
    toc: yes
    toc_float: no
    toc_depth: 2
    theme: paper 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE)
```

## Instructions

- You should submit a R code with the template 'Your-Name.R'
- In your code, before starting a new task, include a block comment with the task number:

```
#*********************************************************************
#                       R Code for Task X
#*********************************************************************
```

## Question 1: Basic R

### Task 1

Create variables `x` equal 5 and `y` equal to 10. Then compute:

- x + y
- x - y
- x*y
- x/y

```{r}

x <- 5
y <- 10

x+y
x-y
x*y
x/y

```

### Task 2

Create a numeric vectors called ```odds``` and ```even``` with odd and even numbers up to 10:

- odds: 1,3,5,7,9
- even: 2,4,6,8,10

```{r T2_solution}

odds <- c(1,3,5,7,9)
even <- c(2,4,6,8,10)

```

### Task 3

Repeat task 2 using the function seq():

If you do not know how to use the seq() function read the documentation acessing by typing ```?seq``` on the console.

```{r T3_solution}

odds <- seq(from = 1, to = 10, by = 2)
even <- seq(from = 2, to = 10, by = 2)

odds
even

```

### Task 4

Using the vectors `odds` and `even` created before, compute:

- odds + even
- odds^2
- even^2

```{r T4_solution}

odds + even
odds^2
even^2

```


### Task 5

Create a matrix called `A` equal to

$$\left[\begin{array}
{rrr}
1 & 2 & 3 \\
4 & 5 & 6 \\
7 & 8 & 9
\end{array}\right]$$


```{r }

A <- matrix(seq(1, 9, 1), nrow = 3, byrow = TRUE)

A
```

### Task 6

Print the 2nd row and 3rd column of matrix `A` created on Task 5.

```{r }

A[2,]

A[,3]
```

### Task 7

Create a matrix called `B` sy summing 3 to each element of matrix `A`.

```{r }

B <- A + 3

B

```

### Task 8

Add matrices `A` and `B` created earlier. What do you get?

```{r }

A + B

```

### Task 9

Create a matrix `C` by concatenating the columns of `A` and `B` using the function `cbind()`.

If you are not sure how to use `cbind()`, check the documentation by typing `?cbind` on the console.

```{r }

C <- cbind(A, B)

C

```

### Task 10

Create a matrix `D` by concatenating the rows of `A` and `B` using the function `rbind()`.

If you are not sure how to use `rbind()`, check the documentation by typing `?rbind` on the console.

```{r }

D <- rbind(A, B)

D

```


## Question 2: Descriptive Statistics

### Task 1

There are built-in data frames examples in R. One of them are called mtcars. What does mtcars look like? Use the function `head()` to see the first 10 rows of this data frame

```{r }

head(iris)

```

### Task 2

Use the command `?iris` to see the documentation for the iris dataset and the function `dim()` to see how many rows and columns are contained in this dataset.

```{r }

?iris
dim(iris)

```

### Task 3

Print the 5th row of the iris dataset

```{r }

iris[5,]

```

### Task 4

Print the column Sepal.Width by using the 'S' sign to access.

```{r }

iris$Sepal.Width

```

### Task 5

Make a histogram of `Sepal.Length` in iris using the function `hist()` (hint: use the ‘$’ sign to access the variable).

```{r }

hist(iris$Sepal.Length)

```

### Task 6

Redo the histogram from Task 5 changing the Title to 'Histogram of Sepal Length', the horizontal axis to 'Sepal Length' and the number of breaks to 20

```{r }

hist(iris$Sepal.Length,
     main = 'Histogram of Sepal Length',
     xlab = 'Sepal Length',
     breaks = 20)

```

### Task 7

Make a boxplot of Sepal.Width in iris using the function `boxplot()`.

```{r }

boxplot(iris$Sepal.Width)

```

### Task 8

Make a boxplot of Sepal.Width in iris using the function `boxplot()` for each type of `Species`. 

```{r }

boxplot(iris$Sepal.Width ~ iris$Species)

```

### Task 9

Make a Scatter plot of `Petal.Length` by `Petal.Width`

```{r }

plot(iris$Petal.Length ~ iris$Petal.Width)

```

### Task 10

Repeat the plot from previous task changing the type of points using `pch=` option.

```{r }

plot(iris$Petal.Length ~ iris$Petal.Width, pch = 16)

```
