---
title: "Laboratory 2 - Introduction to tidyverse"
author: Tamy Tsujimoto
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output: 
  html_document:
    toc: yes
    toc_float: no
    toc_depth: 2
    theme: paper 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE)
```

# Instructions

- You should submit a R code with the template 'Your-Name.R'
- In your code, before starting a new task, include a block comment with the task number:

```{r}
#*********************************************************************
#                       R Code for Task X
#*********************************************************************
```

- For Tasks 1 to 9 you will be using the `student-mat` dataset from [UCI](https://archive.ics.uci.edu/ml/datasets/Student+Performance#).

# Question 1: R Packages and data import

## Task 1

If you have not done so, install the `tidyverse` package using the `install.package` function.

```{r eval = FALSE}
install.packages("tidyverse")
```

## Task 2

Load the `tidyverse` package using the `library()` function

```{r}
library(tidyverse)
```

## Task 3

Import the `student-mat.csvt` using `read.csv()` with option `sep = ';'` and use the function `glimpse()` to check the dataset 

```{r}

math <- read.csv('data/student/student-mat.csv', header = TRUE, sep = ';')
glimpse(math)

```

# Question 2: Introduction to tidyverse

## Task 4

Create a new dataset containing only the following variables:

- sex: student's sex (binary: 'F' - female or 'M' - male) 
- age: student's age (numeric: from 15 to 22)
- address: student's home address type (binary: 'U' - urban or 'R' - rural) 
- traveltime: home to school travel time (numeric: 1 - <15 min., 2 - 15 to 30 min., 3 - 30 min. to 1 hour, or 4 - >1 hour) 
- G1: first period grade (numeric: from 0 to 20) 
- G2: second period grade (numeric: from 0 to 20) 
- G3: final grade (numeric: from 0 to 20, output target)

```{r}

math <-
  math %>% 
  select(sex, age, address, traveltime, G1, G2, G3)

glimpse(math)

```

## Task 5

Using the dataset created in Task 4, filter the female students with age 18 or older from rural area.

```{r}

math %>% 
  filter(sex == 'F', age >= 18, address == 'R')

```

## Task 6

Sort the filtered observations from Task 5 by final grade.

```{r}

math %>% 
  filter(sex == 'F', age >= 18, address == 'R') %>% 
  arrange(G3)

```

## Task 7

Create a new variable `G_avg` by averaging the variables `G1`, `G2`, `G3` for each student.
Use the `glimpse()` function to check your result.

```{r}

math <-
  math %>% 
  mutate(G_avg = (G1+G2+G3)/3) 

glimpse(math)

```

## Task 8

Create a new variable `age_cat` grouping the variable `age` into:

- '< 18'
- '18-20'
- '> 20'

Use the `glimpse()` function to check your result.

```{r}

math <-
  math %>% 
  mutate(age_cat = ifelse(age < 18, '< 18', 
                          ifelse(age <= 20, '18-20', '> 20'))) 

glimpse(math)

```

## Task 9

Compute the number of observations, mean and standard deviation of `G_avg` for each level of `age_cat` using `group_by()` and `summarise()`

```{r}

math %>% 
  group_by(age_cat) %>% 
  summarise(n = n(),
            mean = mean(G_avg),
            sd = sd(G_avg))
```

## Task 10

Use the `gather()` function to create a dataset `math_long` reshaping the `math` dataset into a long format containing three observations per student, the variable `grade` containing the numeric value of all three grades and the variable `period` containing the period of the grade (G1 for first, G2 second, and G3 for final). 

1) Create a variable `id` using `mutate(id = row_number())`
2) Use the function `gather` with appropriate options
3) Sort by `id`

```{r}

math_long <- 
  math %>% 
  mutate(id = row_number()) %>% 
  gather(key = 'period',
         value = 'grade',
         c(G1, G2, G3)) %>% 
  arrange(id)
  

glimpse(math_long)

```

## Task 11

Use the `spread()` function to create a dataset `math_wide` reshaping the `math_long` dataset into a wide format containing one observations per student and the original variable names `G1`, `G2`, `G3` 

1) Use the function `gather` with appropriate options
2) Sort by `id`

```{r}

math_wide <- 
  math_long %>% 
  spread(key = 'period',
         value = 'grade') %>% 
  arrange(id)

glimpse(math_wide)

```

# Question 3: Merging and datasets

For Tasks 12 and 13 you will be using the built in dataset `mtcars`.
First, run the following code to create the toy datasets to be used in the tasks:

```{r}

mtcars1 <-
  mtcars %>% 
  mutate(id = rownames(mtcars)) %>%
  filter(cyl %in% c(6,8)) %>% 
  select(id, mpg:drat)

mtcars2 <-
  mtcars %>% 
  mutate(id = rownames(mtcars)) %>% 
  select(id, wt:carb)

```

## Task 12

Create a dataset called `mtcars_t10` by merging `mtcars1` and `mtcars2` using full join:

```{r}

mtcars_t10 <-
  mtcars1 %>% 
  full_join(mtcars2, by = 'id')

glimpse(mtcars_t10)

```

## Task 13

Create a dataset called `mtcars_t11` by merging `mtcars1` and `mtcars2` using inner join:

```{r}

mtcars_t11 <-
  mtcars1 %>% 
  inner_join(mtcars2, by = 'id')

glimpse(mtcars_t11)

```



