---
title: "Laboratory 3 - Visualization with ggplot"
author: Tamy Tsujimoto
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output: 
  html_document:
    toc: yes
    toc_float: no
    toc_depth: 2
    theme: paper 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, include = F, eval = F)
```

# Instructions

- You should submit a R code with the template 'Your-Name.R'
- In your code, before starting a new task, include a block comment with the task number:

```{r}
#*********************************************************************
#                       R Code for Task X
#*********************************************************************
```

- For Tasks 1 to 9 you will be using the built-in dataset `diamonds` from `ggplot2`

# Question 1: tidyverse and ggplot

## Task 1

Load the packages `tidyverse` and `ggplot2` using the `library()` function

```{r}
library(tidyverse)
library(ggplot2)
```

## Task 2

Access and read the documentation of the `diamonds` dataset using `?diamonds` and use the `glimpse` function to check the dataset.

```{r}

?diamonds
glimpse(diamonds)

```

## Task 3

Create a scatter plot of price in US dollars (`price`) by weight of the diamond (`carat`) using `geom_point`.

```{r}

ggplot(diamonds) +
  geom_point(aes(x = carat, y = price)) 

```

## Task 4

Repeat the plot from Task 3, coloring the points according to clarity of the diamond (`clarity`).

```{r}

ggplot(diamonds) +
  geom_point(aes(x = carat, y = price, color = clarity))

```

## Task 5

Construct a histogram of the prices in US dollars using `geom_histogram` using the options `color = "black"` and `fill = "white"`

```{r}

ggplot(diamonds) +
  geom_histogram(aes(x = price), color = "black", fill = "white")

```

## Task 6

Construct box-plots of price for each diamond color using `geom_boxplot`

```{r}

ggplot(diamonds) +
  geom_boxplot(aes(x = color, y = price))

```

## Task 7

Construct bar plot for the frequencies of quality of the cut (`cut`) using `geom_bar(stat="identity")`.
Customize your plot using `color` and `fill` options.

```{r}

diamonds %>% 
  group_by(cut) %>% 
  summarise(n=n()) %>% 
  ggplot() +
  geom_bar(aes(x = cut, y = n), stat = "identity", color = "red", fill = "pink")

```

## Task 8

Create a new variable `cut_group` by grouping the categories of the variable `cut` as following:

- Fair
- Good/Very Good
- Premium/Ideal

and construct a bar plot of `clarity` for each level of `cut_group` using `facet_wrap`

```{r}

diamonds %>% 
  mutate(cut_group = ifelse(cut %in% c("Good", "Very Good"), "Good/Very Good", 
                                       ifelse(cut %in% c("Premium", "Ideal"), "Premium/Ideal", "Fair"))) %>%
  group_by(cut_group, clarity) %>% 
  summarise(n=n()) %>% 
  ggplot() +
  geom_bar(aes(x = clarity, y = n), stat = "identity") +
  facet_wrap(~ cut_group)

```

