---
title: "Laboratory 4 - Linear Models"
author: Tamy Tsujimoto
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output: 
  html_document:
    toc: yes
    toc_float: no
    toc_depth: 2
    theme: paper 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, include = F, eval = F)
```

# Instructions

- You should submit a R code with the template 'Your-Name.R'
- In your code, before starting a new task, include a block comment with the task number:

```{r}
#*********************************************************************
#                       R Code for Task X
#*********************************************************************
```

- For this lab you will be using the built-in dataset `marketing` from `datarium` package.

## Task 1

If you have not done so, install the `datarium` package using the `install.package` function.

```{r eval = FALSE}
install.packages("datarium")
```

## Task 2

Load the packages `tidyverse`, `ggplot2`, `datarium` and `broom` packages using the `library()` function

```{r}
library(tidyverse)
library(ggplot2)
library(datarium)
library(broom)
```

## Task 3

Access and read the documentation of the `marketing` dataset using `?marketing` and use the `glimpse` function to check the dataset.

```{r}

?diamonds
glimpse(marketing)

```

## Task 4

Construct scatter plots of sales by each of the advertising medias using `ggplot` package.

Include a loess curve and comment the trend in each one of the three plots.

```{r}

marketing %>% 
  ggplot(aes(x = youtube, y = sales)) +
  geom_point() +
  geom_smooth(method = "loess")

marketing %>% 
  ggplot(aes(x = facebook, y = sales)) +
  geom_point() +
  geom_smooth(method = "loess")

marketing %>% 
  ggplot(aes(x = newspaper, y = sales)) +
  geom_point() +
  geom_smooth(method = "loess")

```

## Task 5

Fit a linear model for sales with main effects for each of the three advertising media.

Print the results of the model using the `summary()` function and interpret the coefficients from the model.

```{r}

fit <- lm(sales ~ facebook + youtube + newspaper, data = marketing)
summary(fit)

```

## Task 6

Construct the 95% confidence interval for the coefficients from the fitted model in Task 5.

Interpret the results.

```{r}

confint(fit)

```

## Task 7

Assess the model fit and interpret the results.

```{r}

# add model diagnostics to our training data
model_results <- augment(fit, marketing)

ggplot(model_results, aes(.fitted, .std.resid)) +
  geom_hline(yintercept = 0) +
  geom_point() + geom_smooth(se = FALSE) +
  theme_bw(base_size = 20)

ggplot(model_results, aes(sample = .resid)) + 
  stat_qq() + stat_qq_line() +
  theme_bw(base_size = 20)

ggplot(model_results, aes(seq_along(.cooksd), .cooksd)) +
  geom_bar(stat = "identity", position = "identity") +
  theme_bw(base_size = 20)

ggplot(model_results, aes(.hat, .std.resid)) +
  geom_point(aes(size = .cooksd), na.rm = TRUE)+
  theme_bw(base_size = 20)

```

## Task 8

Summarize the results from this analysis in one paragraph.


