---
title: "Laboratory 5 - Introduction to Survival Analysis"
author: Tamy Tsujimoto and Victor Ritter
date: "`r format(Sys.time(), '%m/%d/%Y')`"
output: 
  html_document:
    toc: yes
    toc_float: no
    toc_depth: 2
    theme: paper 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, include = F, eval = F)
```

# Instructions

- You should submit a R code with the template 'Your-Name.R'
- In your code, before starting a new task, include a block comment with the task number:

```{r}
#*********************************************************************
#                       R Code for Task X
#*********************************************************************
```

- For this lab you will be using the built-in dataset `ovarian` from `survival` package.

## Task 1

If you have not done so, install the `survival` and `survminer` packages using the `install.package` function.

```{r eval = FALSE}
install.packages("survival")
install.packages("survminer")
```

## Task 2

Load the packages `tidyverse`, `survival` and `survminer` packages using the `library()` function

```{r}
library(tidyverse)
library(survival)
library(survminer)
```

## Task 3

Access and read the documentation of the `ovarian` dataset using `?ovarian` and use the `glimpse` and `summary` functions to check the dataset.

```{r}

?ovarian
glimpse(ovarian)
summary(ovarian)

```

## Task 4

Construct the Kaplan-Meier curve for the entire sample using the `survfit()` function.
Plot the curve using `plot()` and `ggsurvplot()` functions to compare.

```{r}

fit <- survfit(Surv(futime, fustat) ~ 1, data = ovarian)

# Using the base R
plot(fit, xlab = "Days",
     ylab = "Overall survival probability")

# Using surviminer
ggsurvplot(fit, data = ovarian, xlab = "Days",
           ylab = "Overall survival probability")

```

## Task 5

Construct the dataset `ovarian2` containing the categorized variable `age_cat` created from `age` according to the following:

- age_cat: 
    + 1: <= 60
    + 2: > 60
    
Check the resulting dataset using the `glimpse()` function
    
```{r}

ovarian2 <-
  ovarian %>% 
  mutate(age_cat = ifelse(age <= 60, 1, 2))

glimpse(ovarian2)

```

## Task 6

Using the `ovarian2` dataset from Task 5, construct the Kaplan-Meier curves with risk tables according to each of the following variables:

- rx:	treatment group 
- resid.ds:	residual disease present (1=no,2=yes)
- ecog.ps:	ECOG performance status (1=Fully active, 2=Restricted)
- age_cat: age in years (1= <= 50, 2= >50)

Comment the KM curves.

```{r}

f1 <- survfit(Surv(futime, fustat) ~ rx, data = ovarian2)
f2 <- survfit(Surv(futime, fustat) ~ resid.ds, data = ovarian2)
f3 <- survfit(Surv(futime, fustat) ~ ecog.ps, data = ovarian2)
f4 <- survfit(Surv(futime, fustat) ~ age_cat, data = ovarian2)

# Using surviminer
ggsurvplot(f1, data = ovarian2, 
           risk.table = TRUE, 
           pval = TRUE,
           xlab = "Days",
           ylab = "Overall survival probability")

ggsurvplot(f2, data = ovarian2,            
           risk.table = TRUE, 
           pval = TRUE,
           xlab = "Days",
           ylab = "Overall survival probability")

ggsurvplot(f3, data = ovarian2, 
           risk.table = TRUE, 
           pval = TRUE,
           xlab = "Days",
           ylab = "Overall survival probability")

ggsurvplot(f4, data = ovarian2, 
           risk.table = TRUE, 
           pval = TRUE,
           xlab = "Days",
           ylab = "Overall survival probability")

```


## Task 7

Using the `ovarian2` dataset from Task 5, perform the log-rank test for each of the following variables:

- rx:	treatment group 
- resid.ds:	residual disease present (1=no,2=yes)
- ecog.ps:	ECOG performance status (1=Fully active, 2=Restricted)
- age_cat: age in years (1= <= 50, 2= >50)

Interpret the results from the log-rank tests.

```{r}

survdiff(Surv(futime, fustat) ~ rx, data = ovarian2)
survdiff(Surv(futime, fustat) ~ resid.ds, data = ovarian2)
survdiff(Surv(futime, fustat) ~ ecog.ps, data = ovarian2)
survdiff(Surv(futime, fustat) ~ age_cat, data = ovarian2)

```

## Task 8

Fit a Cox proportional hazards model including `rx`, `resid.ds`, `ecog.ps` and `age` as covariates. Obs: Do not forget to include the first three variables as factors.

Print the results using the `summary()` function and interpret the results.

```{r}

fit.cox <- coxph(Surv(futime, fustat) ~ factor(rx) + factor(resid.ds) + factor(ecog.ps) + age, 
                 data = ovarian)

summary(fit.cox)
```

## Task 9

Perform the diagnostics to assess if the proportionality assumption from the fitted model from task 8.

Interpret the results.

```{r}

ph <- cox.zph(fit.cox )
ph

ggcoxzph(ph)

```

## Task 10

Summarize the results from this analysis in one paragraph.
