---
title: "Biostatistics Module"
subtitle: "R Lab"
author: "Victor Ritter"
institute: "UNC-CH"
date: "Jan 2020"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      beforeInit: "macros.js"
      # highlightStyle: github
      # highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = ">",
                      cache = TRUE,
                      cache.path = "./cache/")
options(htmltools.dir.version = FALSE,
        dplyr.width = 60,
        dplyr.print_min = 5,
        dplyr.print_max = 5)

library(tidyverse)
library(caret)
library(ResourceSelection)
library(broom)
library(pROC)
```

class: center, middle, inverse
# Part I
# Logistic Regression

---
## Data

- Analysis based on the Coursera course “Introduction to Logistic Regression” by Prof. Stanley Lemeshow (OSU-USA)

- _Applied Logistic Regression_ by Hosmer & Lemeshow (2013)

--

- Low birth weight study data aiming to identify risk factors associated with giving birth to a low birth weight baby in the population being served by the medical centre where the data were collected.

- We'll be restricted to
  - `LOW`: Low Birth Weight (1=Birth Weight < 2.5kg; 0=otherwise)
  - `AGE`: moms age
  - `LWT`: baby weight at last menstrual period
  - `RACE`: mom's race (1 = White, 2 = Black, 3 = Other)

???

Ohio State Uni. College of PH

---
## Data - descriptive analysis

```{r}
LOWBWT = read.table("./data/LOWBWT.txt", header = TRUE) %>% 
  mutate_at(vars(LOW, RACE), as.factor)
fq <- LOWBWT %>% {
  table(.$LOW, .$RACE)
}
margin.table(fq, 1)
prop.table(fq, margin = 1)
```

---
## Data - descriptive analysis

```{r, echo=T, message=F, warning=F, fig.width=7, fig.height=6, fig.align='center'}
# Overlaid histograms
ggplot(LOWBWT, aes(x = AGE, fill = LOW)) +
  geom_density(alpha=.5)
```

---
## Data - descriptive analysis

```{r, echo=T, message=F, warning=F, fig.width=7, fig.height=6, fig.align='center'}
ggplot(LOWBWT, aes(x = LWT, fill = LOW)) +
  geom_density(alpha=.5)
```

---
## Model - fit

```{r}
fit = glm(LOW ~ AGE + LWT + RACE,
          family = binomial(link = "logit"), data = LOWBWT)
fit
```

---
## Model - fit

```{r}
summary(fit)
```

---
## Model - analysis of deviance table

```{r}
anova(fit, test = "Chisq")
```

---
## Model - analysis of deviance table

```{r}
car::Anova(fit)
```

---
## Model - analysis of deviance table

```{r}
fit1 <- update(fit, . ~ AGE + LWT * RACE)
anova(fit1, fit, test = "LR")
```

---
## Model - goodness-of-fit

- Classic GOF statistics

```{r}
# Deviance
g2 <- sum(residuals(fit, type = "deviance")^2)
g2
1 - pchisq(g2, df = fit$df.residual)

# Person 
p <- sum(residuals(fit, type = "pearson")^2)
p
1 - pchisq(p, df = fit$df.residual)

```

---
## Model - goodness-of-fit

- Hosmer-Lemeshow test

```{r}
library(ResourceSelection) 
hoslem.test(fit$y, fitted(fit), g = 10)
```

---
## Model - overdispersion

```{r}
summary(update(fit, . ~ ., family = quasibinomial))
```

---
## Model - residuals

```{r}
library(broom)
fit_resid <- augment(fit) %>% 
  mutate(index = 1:n())
fit_resid
```

---
## Model - residuals

```{r, echo=T, message=F, warning=F, fig.width=7, fig.height=6, fig.align='center'}
ggplot(fit_resid, aes(index, .std.resid, color = LOW)) + 
  geom_point(size = 2)
```

???
Those standardized residuals that exceed 3 represent possible outliers and may deserve closer attention.

---
## Model - residuals

```{r, echo=T, message=F, warning=F, fig.width=7, fig.height=6, fig.align='center'}
ggplot(fit_resid, aes(index, .std.resid)) + 
  geom_point(size = 2) + facet_wrap(~ LOW, scales = "free")
```

---
## Model - residuals

```{r, echo=T, message=F, warning=F, fig.width=7, fig.height=5, fig.align='center'}
ggplot(fit_resid, aes(x = .fitted, y = .cooksd,
                      ymin = 0, ymax = .cooksd)) + 
  geom_point(size = 2) + geom_linerange() +
  geom_label(aes(x = .fitted, y = .cooksd, label = index),
             data = top_n(fit_resid, 5, .cooksd))
```

---
## Model - agreement

```{r}
pred <- predict(fit, type = "response")
pred[1:20]

bin_pred <- as.numeric(pred > 0.5) %>% factor()
table(bin_pred, LOWBWT$LOW)
```

---
## Model - agreement

```{r}
library(caret)
confusionMatrix(data = bin_pred, reference = LOWBWT$LOW)
```

---
## Model - agreement

```{r}
library(pROC)
roc <- roc(LOWBWT$LOW ~ pred)
roc
```

---
## Model - agreement

```{r, echo=T, message=F, warning=F, fig.width=7, fig.height=6, fig.align='center'}
plot(roc, print.auc = T)
```
