---
title: "Biostatistics Module"
subtitle: "R Lab"
author: "Victor Ritter"
institute: "UNC-CH"
date: "31/01/2020"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      beforeInit: "macros.js"
      # highlightStyle: github
      # highlightLines: true
      countIncrementalSlides: false
---

```{css}
.tiny .remark-code {
  font-size: 70% !important;
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment = ">",
                      cache = FALSE,
                      cache.path = "./cache/")
options(htmltools.dir.version = FALSE,
        dplyr.width = 60,
        dplyr.print_min = 5,
        dplyr.print_max = 5)

library(lme4)
library(lmerTest)
library(broom.mixed)
library(geepack)
library(tidyverse)
```
---

class: center, middle, inverse
# Part I
# Analysis of repeated measures <br/> in R

---
## GEE - basics

- population-averaged estimates and their standard errors

- The responses $Y_1, Y_2, \dots, Y_n$ are correlated or clustered

- There is a linear relationship between the covariates and a transformation of the mean response (i.e. link function)

- Within-subject covariance has some structure (“working covariance”):

  - independence (observations over time are independent)
  
  - exchangeable (all observations over time have the same correlation)
  
  - AR(1) (correlation decreases as a power of how many time-points apart two observations are)
  
  - unstructured (correlation between all time-points may be different)


---
## GEE - basics

- misspecified covariance: use “sandwich estimator” for robustness

  - idea behind is to use the empirical residuals to approximate the underlying covariance (iteratively reweighted least squares)
  
- Why bother specifying the working covariance to begin with?

  - Statistical efficiency
  
  - Sandwich robustness is a large-sample property

- Should we use sandwich all the time? No, it is problematic if

  - The number of independent subjects is much smaller than the number of repeated measures
  
  - The design is unbalanced – the number of repeated measures differs across individuals

---
## GEE - package and data

- **geepack** package

- **dietox** data available
  
  - Growth curves of pigs in a 3x3 factorial experiment
  
  - interest in modeling pig's weight as a function of time, vitamin E, and Copper supplement dosage
  
```{r}
library(geepack)
library(tidyverse)
as_tibble(dietox)
```
  
---
## GEE - package and data

```{r}
dietox %>%
  group_by(Pig) %>%
  count() %>%
  arrange(n) %>%
  print(n = 10)
```

---
## GEE - descriptive

Marginal covariance

```{r}
pig_wide <- dietox %>%
  dplyr::select(Pig, Weight, Time) %>% 
  spread(Time, Weight) %>% 
  as_tibble()
pig_wide
```

---
## GEE - descriptive

```{r, fig.width=7, fig.height=5, fig.align='center'}
dietox %>%
  ggplot(aes(x = Time, y = Weight, group = Pig)) +
  geom_point() + geom_line() +
  stat_summary(fun.y=mean, aes(group=1),
               geom="line", colour="red", size = 1.5)
```

---
## GEE - descriptive

```{r, fig.width=7, fig.height=5, fig.align='center'}
dietox %>%
  ggplot(aes(x = Time, y = Weight, group = Pig, color = Cu)) +
  geom_point() +
  geom_line()
```

---
## GEE - descriptive

Marginal covariance

```{r}
cor(pig_wide[,-1], use = "pairwise.complete.obs") %>% round(2)
```
  
  
---
## GEE - model fit

```{r echo=FALSE}
mf <- formula(Weight ~ Time + Evit + Cu)
geeAr1 <- geeglm(mf, id=Pig, data=dietox,
                 family=gaussian, corstr="ar1")
```
.tiny[
```{r}
summary(geeAr1)
```
]

---
## GEE - model fit

Independent correlation structure

```{r}
geeInd <- geeglm(mf, id=Pig, data=dietox,
                 family=gaussian, corstr="ind")
```
.tiny[
```{r}
summary(geeInd)
```
]

---
## GEE - model fit

- On the choice of the correlation structure

```{r}
MuMIn::QIC(geeAr1)
MuMIn::QICu(geeAr1)
MuMIn::QIC(geeInd)
MuMIn::QICu(geeInd)
```

---
## Linear Mixed Models - basics

- Subject-specific or cluster-specific model of correlated/clustered data

- Basic premise is that there is natural heterogeneity across individuals in the
study population that is the result of unobserved covariates; random effects
account for the unobserved covariates

- Package **lme4** fit Linear (**`lmer()`**) and Generalized Linear (**`glmer()`**) Mixed models

- Model formula is of the form:

.center[y ~ fixed effects + (random effect | cluster)]

  - Random intercept model: `(1 | id)`
  
  - Random slope model: `(0 + x | id)`
  
  - Random slope and intercept model: `(1 + x | id)` or just `(x | id)`


---
## Linear Mixed Models - model fit

Modeling **individual means** with random intercept

```{r}
mm1 <- lmer(Weight ~ Time + Evit + Cu +
              (1|Pig), data=dietox)
```
```
summary(mm1)
```

---
**summary(mm1)**

.tiny[
```{r echo=FALSE}
summary(mm1)
```
]

---
## Linear Mixed Models - model fit

```{r}
broom.mixed::glance(mm1)
broom.mixed::tidy(mm1, conf.int = T)
```

---
## Linear Mixed Models - model fit

Modeling **individual means** with random intercept

```{r}
mm2 <- lmer(Weight ~ Time + Evit + Cu +
              (1+Time|Pig), data=dietox)
```
```
summary(mm2)
```

---
## LMM - testing fixed-effects

Use ML estimates, i.e. **REML = FALSE**

```{r}
fix1 <- lmer(Weight ~ Time + Evit + Cu +
              (1|Pig), data=dietox, REML = F)
fix2 <- lmer(Weight ~ Time + Evit +
              (1|Pig), data=dietox, REML = F)

anova(fix1, fix2, refit = FALSE)
```

---
## LMM - testing random-effects

Use ML estimates, i.e. **REML = TRUE**

```{r}
rd1 <- lmer(Weight ~ Time + Evit +
              (1|Pig), data=dietox, REML = T)
rd2 <- lmer(Weight ~ Time + Evit +
              (1 + Time|Pig), data=dietox, REML = T)

anova(rd1, rd2, refit = FALSE)
```

---
## LMM - testing random-effects

Except when testing if **ANY** random-effect is necessary

```{r}
rd1 <- lmer(Weight ~ Time + Evit +
              (1|Pig), data=dietox, REML = F)
rd2 <- lm(Weight ~ Time + Evit, data=dietox)

dev1 <- -2*logLik(rd1)
dev0 <- -2*logLik(rd2)
devdiff <- as.numeric(dev0-dev1); devdiff
dfdiff <- attr(dev1,"df") - attr(dev0,"df"); dfdiff
cat('Chi-square =', devdiff, '(df=', dfdiff,'), p =', 
  pchisq(devdiff, dfdiff, lower.tail=FALSE))
```













